---
title: "Supplementary Material for Effects of bariatric surgery and weight loss on resting state functional connectivity"
author:
- Hannah Sophie Heinrichs* [1]
- Frauke Beyer* [1][2]
- Kristin Prehn
- Jürgen Ordemann 
- Agnes Floel
- A. Veronica Witte [1][2][3]
date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3
  Day Clinic for Cognitive Neurology, University Clinic Leipzig
output:
  bookdown::pdf_book:
    fig_caption: yes
    always_allow_html: true
    number_sections: yes
    toc: yes
    extra_dependencies: "subfig"
bibliography: "bibliography.bib"
editor_options:
  chunk_output_type: console
---

```{r "load packages", include=FALSE}

# load packages
library(car) # version 3.0.9
library(plyr) # version 1.8.6
library(dplyr) # version 1.0.2
library(tidyr) # version 1.1.2
library(kableExtra) # version 1.2.1
library(ggplot2) # version 3.3.2
library(patchwork) # version 1.0.1; API for sequentially building up a plot (similar to gridExtra and cowplot)
library(cowplot)
library(RColorBrewer) # version 1.1.2
library(psych) # version 2.0.8

#library(lsr)
library(wesanderson) # colour palette
#library(haven)

library(lme4)

library(sjPlot) #only works for HTML output
library(texreg) #works for Rmarkdown
library(knitr) # version 1.30


source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/glmm_stability.r')
source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/diagnostic_fcns.r')

```

```{r "knit setup", include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
options(knitr.table.format = "latex") 
options(knitr.kable.NA = '') # hide NA in tables
# set working directory of knit (RMarkdown) to directory of Rproj
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
setwd(knitr::opts_knit$get("root.dir"))
```


## Interaction effect group by time on aggregated FC (preregistered analysis)
```{r load agg FC data and prepare, echo=FALSE}
agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
for (i in c(1:nrow(agg_FC_DMN_ID))){
  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
}

colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID_tp")
colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID_tp")

final_FC=merge(final, agg_FC_Rew, by="subj.ID_tp", all.x=TRUE)
final_FC=merge(final_FC, agg_FC_DMN, by="subj.ID_tp", all.x=TRUE)
final_FC$subj.ID=as.factor(final_FC$subj.ID)
final_FC$group_factor=relevel(final_FC$group_factor, ref = "KG")

#calculate mean BMI over timepoints
final_FC.meanBMI=tapply(X=final_FC$BMI,
                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
final_FC$mean.BMI=
  final_FC.meanBMI[as.numeric(final_FC$subj.ID)]
final_FC$within.BMI=
  final_FC$BMI-final_FC$mean.BMI


#calculate mean FD over timepoints
final_FC.meanFD=tapply(X=final_FC$logmFD,
                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
final_FC$mean.logmFD=
  final_FC.meanFD[as.numeric(final_FC$subj.ID)]
final_FC$within.logmFD=
  final_FC$logmFD-final_FC$mean.logmFD
```

### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
```{r calc DMN two tp, age and sex,results = 'asis', echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvDMN2tp=anova(R1,R0)
```

```{qa check DMN model age,sex, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex, data=final_FC)
vif_res=vif(test_vif)
```

```{r DMN2tp results age, sex, caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)

```{r calc DMN two tp, age and sex and mFD,results = 'asis', echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvDMN2tp=anova(R1,R0)
```

```{qa check DMN model age,sex, mFD, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
vif_res=vif(test_vif)
```

```{r DMN2tp results age sex mFD, caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "logmFD", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Aggregated DMN for three timepoints, adjusted for age, sex, mean FD (Model 1b)
```{r calc DMN three tp, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
anvDMN3tp=anova(R1,R0)
```

```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r DMN conn over three timepoints, echo=FALSE}
fig_DMNconn<-mk_figDMNdescr(final_FC)
fig_DMNconn
```

### Reward network for two time points, adjusted for age and sex (Model 1a)
```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvRew2tp=anova(R1,R0)
```
```{qa check Rew model age,sex, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary
```

```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Reward network for two timepoints, adjusted for age, sex, mean FD (Model 1b)
```{r calc rew two tp adjusted, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvRew2tpadj=anova(R1,R0)
```

```{qa check Rew model age,sex,mFD, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
vif_res=vif(test_vif)
```

```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","logmFD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Reward network for three timepoints, adjusted for age, sex, mean FD (Model 1b)
```{r calc rew three tp, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
anvRew3tp=anova(R1,R0)
```
```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```
```{r Rew conn over three timepoints, echo=FALSE}
fig_Rewconn<-mk_figRewdescr(final_FC)
fig_Rewconn
```
