---
title: "Supplementary Material for Effects of bariatric surgery and weight loss on resting state functional connectivity"
author:
- Hannah Sophie Heinrichs* [1]
- Frauke Beyer* [1][2]
- Kristin Prehn
- Jürgen Ordemann 
- Agnes Floel
- A. Veronica Witte [1][2][3]
date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3
  Day Clinic for Cognitive Neurology, University Clinic Leipzig
output:
  bookdown::pdf_book:
    fig_caption: yes
    always_allow_html: true
    number_sections: yes
    toc: no
    extra_dependencies: "subfig"
bibliography: "bibliography.bib"
editor_options:
  chunk_output_type: console
---

```{r "load packages", include=FALSE}

# load packages
library(car) # version 3.0.9
library(plyr) # version 1.8.6
library(dplyr) # version 1.0.2
library(tidyr) # version 1.1.2
library(kableExtra) # version 1.2.1
library(ggplot2) # version 3.3.2
library(patchwork) # version 1.0.1; API for sequentially building up a plot (similar to gridExtra and cowplot)
library(cowplot)
library(RColorBrewer) # version 1.1.2
library(psych) # version 2.0.8

#library(lsr)
#library(haven)
library(wesanderson) # colour palette

library(lme4)

library(sjPlot) #only works for HTML output
library(texreg) #works for Rmarkdown
library(knitr) # version 1.30

source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/glmm_stability.r')
source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/diagnostic_fcns.r')
```

```{r "knit setup", include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
options(knitr.table.format = "latex")
options(knitr.kable.NA = ' ') # hide NA in tables
# set working directory of knit (RMarkdown) to directory of Rproj
```
\newpage

# Methods
## Sample
```{r "tableSample", echo=FALSE}

tableSample <- readRDS(file = "../report/tab/tableSample.rds")

tab <-
  knitr::kable(
    tableSample,
    row.names = TRUE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = "Distribution of data points at months after intervention")
tab
```

```{r tableDescr, include=FALSE}
# table for descriptive statistics
tableDescr <- readRDS(file = "../report/tab/tableDescr.rds")

knitr::kable(
    tableDescr,
    row.names = FALSE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = "BMI and mFD of participants acquired in each condition and time point")
```

## Design
```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap = c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
DesignMatricesList <- readRDS(file = "../report/fig/DesignMatricesList.rds")
DesignMatricesList$plotA
DesignMatricesList$plotB
#gridExtra::grid.arrange(plotA, plotB, nrow=1)
```

# Results
## Quality Control
```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%", eval=TRUE}
tsnr <- "../report/tsnr.jpeg"
knitr::include_graphics(tsnr)
```

## Interaction effect group by time on aggregated FC (preregistered analysis)
```{r load agg FC data and prepare, echo=FALSE}
final_FC <- read.csv("../report/final_FC.csv")
```

### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
```{r calc DMN two tp, age and sex,results = 'asis', echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvDMN2tp=anova(R1,R0)
```

```{r qa check DMN model age,sex, eval=FALSE}
diag=diagnostics.plot(R0)
randefplot=ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex, data=final_FC)
vif_res=vif(test_vif)
```

```{r DMN2tp results age, sex, caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)

```{r calc DMN two tp, age and sex and mFD,results = 'asis', echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvDMN2tp=anova(R1,R0)
```

```{r qa check DMN model age,sex, mFD, eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
vif_res=vif(test_vif)
```

```{r DMN2tp results age sex mFD, caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "logmFD", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Aggregated DMN for three timepoints, adjusted for age, sex, mean FD (Model 1b)
```{r calc DMN three tp, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
anvDMN3tp=anova(R1,R0)
```

```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r DMN conn over three timepoints, echo=FALSE}
fig_DMNconn <- readRDS(file = "../report/fig/figDMNconn.rds")
fig_DMNconn
```

### Reward network for two time points, adjusted for age and sex (Model 1a)
```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvRew2tp=anova(R1,R0)
```
```{r qa check Rew model age,sex, eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary
```

```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Reward network for two timepoints, adjusted for age, sex, mean FD (Model 1b)
```{r calc rew two tp adjusted, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvRew2tpadj=anova(R1,R0)
```

```{r qa check Rew model age,sex,mFD,eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
vif_res=vif(test_vif)
```

```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","logmFD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Reward network for three timepoints, adjusted for age, sex, mean FD (Model 1b)
```{r calc rew three tp, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
anvRew3tp=anova(R1,R0)
```
```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```
```{r Rew conn over three timepoints, echo=FALSE}
fig_Rewconn <- readRDS(file = "../report/fig/figRewconn.rds")
fig_Rewconn
```

## Effects of average and change BMI on FC
### DMN
#### Model 2a and 2b (adjusting for mFD)
```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]

R11 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
R12a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
anvmeanBMI_adjmFD=anova(R11,R12a)
```

```{r DMN_BMIFD results, caption="Model 2a and 2b for aggregated DMN FC", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R1,R11), custom.coef.names=c("Intercept", "average BMI", "change BMI", "Age", "Sex","logmFD"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```
#### Model 2c (adjusting for average and change in mFD)
```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
anvmeanFD=anova(R1,R03)
anvwithinFD=anova(R1,R03)

p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
```

```{r DMN_BMI_splitFD results, caption="Model 2c for aggregated DMN FC", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R1), custom.coef.names=c("Intercept", "average BMI", "change BMI", "average mFD", "change mFD", "Age", "Sex"), custom.model.names = "R1", float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```
\newpage
```{r, include = FALSE}
library(dplyr)
DIR_ANALYSIS <- "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis"
df <- file.path(DIR_ANALYSIS,"noExclFD/result_report.txt")
AnatomyResults <- read.table(df, sep="\t",col.names = c("model","dir","k"),stringsAsFactors = FALSE)

  
txtList <- list()

for (r in 1:nrow(AnatomyResults)) {
  
  file = AnatomyResults[r,"dir"]
  txt_cols <- c("VoxelCount","equals","PercClusterVolumeAssignedTo","in","Hem","Region","PercentOfArea","V8","V9","V10","V11","V12","V13")
  txt <- read.table(trimws(f),sep="\t",fill=TRUE,col.names=txt_cols) # paste0("V", 1:13)
  

  
  txt <- txt[sort(c(which(grepl("Cluster", txt$VoxelCount)),which(grepl("voxel", txt$equals)))),]
  txt <- txt[,c("VoxelCount","PercClusterVolumeAssignedTo", "Hem","Region","PercentOfArea")]
  
  idxClusterStart <- which(grepl("Cluster", txt$VoxelCount))
  
  #txt$VoxelCount <- as.character(txt$VoxelCount)
  txt <- cbind(descr = NA, txt)
  txt$descr[idxClusterStart] <- stringr::str_remove(string = as.character(txt$VoxelCount[idxClusterStart]), pattern = "\\:[^:]*$")
  txt$VoxelCount[idxClusterStart] <- NA
  
  # attach model with  cluster list to list of models
  name_model <- AnatomyResults$model[r]
  txtList[[name_model]] <- txt
}

txt_colname <- kableExtra::linebreak(c("Cluster","Number of voxels\n in cluster","\\% of cluster volume\nassigned", "Hemisphere","Area","\\% of area overlap\nwith cluster"))

TableList <- list()
for (i in 1:length(txtList)){
  TableList[[i]] <-
    txtList[[i]] %>%
    knitr::kable(
      escape = FALSE,     # use font spec of latex with kableExtra
      col.names = txt_colname,
      row.names = FALSE,
      format = "latex",
      booktabs = T,
      linesep = "",      # disable "\\addlinespace" at every 5th line
      #label = name_model[i],
      caption = sprintf("Anatomical labelling for %s",name_model[i]),
    ) %>%
    kableExtra::kable_styling(latex_options = c('scale_down')) #%>% 
    #kableExtra::landscape()
}
```

### Anatomical Labelling for models

```{r, echo=FALSE}

AnatomyResults$model <- read.table(df, sep="\t",col.names = c("model","dir","k"),stringsAsFactors = FALSE)[1]
l <- stringr:: str_split(AnatomyResults$model, '_')
d <- data.frame(matrix(unlist(l), nrow=length(l), byrow=T))
#stringr::str_replace(as.character(d[,2]),"^\\(PCC|NACC)cc$"," (cc)") # , "\\(PCC|NACC)(gsr)" = "(gsr)"

TableList[[1]]
TableList[[2]]
TableList[[3]]
TableList[[4]]
TableList[[5]]
TableList[[6]]
TableList[[7]]
TableList[[8]] 
TableList[[9]]
```