---
title: "Supplementary Material for Effects of bariatric surgery and weight loss on resting state functional connectivity"
author:
- Hannah Sophie Heinrichs* [1]
- Frauke Beyer* [1][2]
- Evelyn Medawar [1]
- Kristin Prehn [3][4]
- Jürgen Ordemann [8][9]
- Agnes Floel [5][6][7]
- A. Veronica Witte [1][2][10]
date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 4 Department of Psychology, Medical School Hamburg, Hamburg \newline 5 Department of Neurology, University of Greifswald, Greifswald \newline 6 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 7 Center for Stroke Research, Charité University Medicine, Berlin \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Zentrum für Adipositas und Metabolische Chirurgie, Vivantes Klinikum Spandau, Berlin \newline 10 Day Clinic for Cognitive Neurology, University Clinic Leipzig
output:
  bookdown::pdf_book:
    fig_caption: yes
    always_allow_html: true
    number_sections: yes
    toc: no
    extra_dependencies: ["float","subfig"]
bibliography: "bibliography.bib"
editor_options:
  chunk_output_type: console
---

```{r "load packages", include=FALSE}
library(car) # version 3.0.9
library(plyr) # version 1.8.6
library(dplyr) # version 1.0.2
library(tidyr) # version 1.1.2
library(kableExtra) # version 1.2.1
library(ggplot2) # version 3.3.2
library(patchwork) # version 1.0.1; API for sequentially building up a plot (similar to gridExtra and cowplot)
library(cowplot)
library(RColorBrewer) # version 1.1.2
library(psych) # version 2.0.8

#library(lsr)
#library(haven)
library(wesanderson) # colour palette

library(lme4)

library(sjPlot) # only works for HTML output
library(texreg) # for works for Rmarkdown
library(knitr) # version 1.30

source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/glmm_stability.r')
source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/diagnostic_fcns.r')
```

```{r "knit setup", include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
options(knitr.table.format = "latex")
options(knitr.kable.NA = ' ') # hide NA in tables
knitr::opts_chunk$set(fig.pos = "!ht", out.extra = "")
# set working directory of knit (RMarkdown) to directory of Rproj
```
\newpage

# Methods
## Sample
```{r "tableSample", echo=FALSE}
tableSample <- readRDS(file = "../report/tab/tableSample.rds")
tab <-
  knitr::kable(
    tableSample,
    row.names = TRUE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = "Distribution of data points at months after intervention")
tab
```

```{r tableDescr, include=FALSE}
# table for descriptive statistics
tableDescr <- readRDS(file = "../report/tab/tableDescr.rds")
knitr::kable(
    tableDescr,
    row.names = FALSE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = "BMI and mFD of participants acquired in each condition and time point")
```


## Quality Control

```{r fighistograms,fig.cap="Histograms of FC-QC correlations for different preprocessing schemes)", out.width="100%", fig.align = "center", eval=FALSE}
tsnr <- "../report/tsnr.pdf"
knitr::include_graphics(tsnr)
```


## Networks
Figure \@ref(fig:figNetworks) shows the DMN and reward network average maps which we used for extracting aggregated FC values. Both networks included the typically described brain regions (i.e. medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network).
Yet, the reward network was less pronounced (overall lower T-values) and less symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see \@ref(fig:figtSNR)).
Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) are on Neurovault.
```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all time points and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$)", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}

networks = c("../report/fig/DMN_thresholded_for_clusterFWE005.pdf",
             "../report/fig/Rew_thresholded_for_clusterFWE005.pdf")
knitr::include_graphics(networks)
```

```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="100%", fig.align = "center", eval=TRUE}
tsnr <- "../report/tsnr.pdf"
knitr::include_graphics(tsnr)
```



## Design
```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap = c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
DesignMatricesList <- readRDS(file = "../report/fig/DesignMatricesList.rds")
DesignMatricesList$plotA
DesignMatricesList$plotB
#gridExtra::grid.arrange(plotA, plotB, nrow=1)
```

## Aggregated FC: confirmatory analysis of an interaction effect of group and time 
```{r load agg FC data and prepare, echo=FALSE}
final_FC <- read.csv("../report/final_FC.csv")
final_FC$group_factor=relevel(final_FC$group_factor, "KG")
final_FC=merge(final_FC,subset(final_FC, tp=="bl",select=c("subj.ID", "BMI")), by='subj.ID')
colnames(final_FC)[27]="BMI.bl"
```


```{r identify outliers in these plots, eval=FALSE}
### SUMMARY: all in all there is no reason to exclude the subjects with outlying values in mean Rew/DMN connectivity
final_FC[final_FC$mean_DMN_conn< 0.1, c("mean_DMN_conn", "mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")]
#ADI003_fu2 & ADI041_fu -> do not have particularly high head motion
#plot(final_FC$mean_DMN_conn, final_FC$logmFD) -> see two left points
#carpet plots look ok, brain maps of DMN look okish, though not pronounced
#ADI003_fu2 & ADI041 fu: FS issues (Final_score:3, corrected)
qa_info=read.csv("/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv")
qa_info[qa_info$subj.ID=="ADI003"&qa_info$tp=="fu2",]
qa_info[qa_info$subj.ID=="ADI041"&qa_info$tp=="fu",]

final_FC[final_FC$mean_Rew_conn>0.3, c("mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")]
plot(final_FC$mean_Rew_conn, final_FC$logmFD)
#-> do not have particularly high head motion (two right points)
#ADI069_bl
#nifti looks ok, carpet plot shows lot of stripes.
qa_info[qa_info$subj.ID=="ADI069"&qa_info$tp=="bl",]
#FS issues, parts of skull included
#ADI111_fu2
#nifti looks ok, some residual carpet plot noise
qa_info[qa_info$subj.ID=="ADI111"&qa_info$tp=="fu2",]
#no FS issues
```


### Reward network 
#### Two time points, without adjustments (Model CA1) 

<b>

```{r calc rew two tp, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvRew2tp=anova(R1,R0)
```

```{r qa check Rew model age,sex, eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary
```

```{r Rew2tp results, caption="Interaction of group and time on reward network connectivity for two time points", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Group BARS",
    "Timepoint x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

#### Two timepoints, adjusted for age, sex and average mFD (Model CA2)

```{r calc rew two tp adjusted 2, fig.cap="Two timepoints, adjusted for age, sex and average mFD (Model CA2)", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvRew2tpadj=anova(R1,R0)
```

```{r qa check Rew model age,sex,mFD,eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
vif_res=vif(test_vif)
```

```{r Rew2tpadj2 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  reorder.coef = c(1, 2, 3, 7, 4, 5, 6),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Group BARS",
    "Age",
    "Sex",
    "av mFD",
    "Timepoint x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

#### Two timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)

```{r calc rew two tp adjusted 3, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvRew2tpadj=anova(R1,R0)
```

```{r qa check Rew model age,sex,mFD,baselineBMI,eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI.bl, data=final_FC)
vif_res=vif(test_vif)
```

```{r Rew2tpadj3 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  reorder.coef = c(1, 2, 3, 8, 4, 5, 6,7),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Group BARS",
    "Age",
    "Sex",
    "av mFD",
    "baseline BMI",
    "Timepoint x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

#### Three time points, without adjustments (Model CA1)

```{r calc rew three tp, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + (1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor +  (1|subj.ID), data=final_FC)
anvRew2tp=anova(R1,R0)
```
```{r qa check Rew model age,sex three timepoints, eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary
```

```{r Rew3tp results, caption="Interaction of group and time on reward network connectivity for two time points", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Timepoint FU2",
    "Group BARS",
    "Timepoint FU x Group", 
    "Timepoint FU2 x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

#### Three timepoints, adjusted for age, sex and average mFD (Model CA2)

```{r calc rew three tp adjusted 2, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + (1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC)
anvRew2tpadj=anova(R1,R0)
```

```{r qa check Rew model age,sex,mFD, three timepoints,eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
vif_res=vif(test_vif)
```

```{r Rew3tpadj 2 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  reorder.coef = c(1, 2, 3, 4, 8, 9, 5, 6, 7),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Timepoint FU2",
    "Group BARS",
    "Age",
    "Sex",
    "av mFD",
    "Timepoint FU x Group", 
    "Timepoint FU2 x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

#### Three timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)

```{r calc rew three tp adjusted 3, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC)
anvRew2tpadj=anova(R1,R0)
```

```{r qa check Rew model age,sex,mFD,baselineBMI, three timepoints,eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI.bl, data=final_FC)
vif_res=vif(test_vif)
```

```{r Rew3tpadj 3 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  reorder.coef = c(1, 2, 3, 4,9,10, 5, 6, 7,8),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Timepoint FU2",
    "Group BARS",
    "Age",
    "Sex",
    "av mFD",
    "baseline BMI",
    "Timepoint FU x Group", 
    "Timepoint FU2 x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

```{r Rew conn over three timepoints, echo=FALSE}
fig_Rewconn <- readRDS(file = "../report/fig/figRewconn.rds")
fig_Rewconn
```


### DMN 
#### Two time points, without adjustments (Model CA1)

```{r calc rew two tp DMN, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvDMN2tp=anova(R1,R0)
```
```{r qa check DMN model age,sex, eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary
```

```{r DMN2tp results, caption="Interaction of group and time on DMN connectivity for two time points", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Group BARS",
    "Timepoint x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

#### Two timepoints, adjusted for age, sex and average mFD (Model CA2)

```{r calc DMN two tp adjusted 2, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvDMN2tpadj=anova(R1,R0)
```

```{r qa check DMN model age,sex,mFD,eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
vif_res=vif(test_vif)
```

```{r DMN2tpadj2 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  reorder.coef = c(1, 2, 3, 7, 4, 5, 6),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Group BARS",
    "Age",
    "Sex",
    "av mFD",
    "Timepoint x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

#### Two timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)

```{r calc DMN two tp adjusted 3, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvDMN2tpadj=anova(R1,R0)
```

```{r qa check DMN model age,sex,mFD,baselineBMI,eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI.bl, data=final_FC)
vif_res=vif(test_vif)
```

```{r DMN2tpadj3 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  reorder.coef = c(1, 2, 3, 8, 4, 5, 6,7),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Group BARS",
    "Age",
    "Sex",
    "av mFD",
    "baseline BMI",
    "Timepoint x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

#### Three time points, without adjustments (Model CA1)

```{r calc DMN three tp, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + (1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor +  (1|subj.ID), data=final_FC)
anvDMN3tp=anova(R1,R0)
```
```{r qa check DMN model age,sex three timepoints, eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary
```

```{r DMN3tp results, caption="Interaction of group and time on DMN connectivity for two time points", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Timepoint FU2",
    "Group BARS",
    "Timepoint FU x Group", 
    "Timepoint FU2 x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

#### Three timepoints, adjusted for age, sex and average mFD (Model CA2)

```{r calc DMN three tp adjusted 2, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + (1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC)
anvDMN3tpadj=anova(R1,R0)
```

```{r qa check DMN model age,sex,mFD, three timepoints,eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
vif_res=vif(test_vif)
```

```{r DMN3tpadj 2 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  reorder.coef = c(1, 2, 3, 4, 8, 9, 5, 6, 7),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Timepoint FU2",
    "Group BARS",
    "Age",
    "Sex",
    "av mFD",
    "Timepoint FU x Group", 
    "Timepoint FU2 x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

#### Three timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)
```{r calc DMN three tp adjusted 3, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC)
anvDMN3tpadj=anova(R1,R0)
```

```{r qa check DMN model age,sex,mFD,baselineBMI, three timepoints,eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI.bl, data=final_FC)
vif_res=vif(test_vif)
```

```{r DMN3tpadj 3 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R0, R1),
  reorder.coef = c(1, 2, 3, 4, 9, 10, 5, 6, 7,8),
  custom.coef.names = c(
    "Intercept (BL, NBARS)",
    "Timepoint FU",
    "Timepoint FU2",
    "Group BARS",
    "Age",
    "Sex",
    "av mFD",
    "baseline BMI",
    "Timepoint FU x Group", 
    "Timepoint FU2 x Group"
  ),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```

```{r DMN conn over three timepoints, echo=FALSE}
fig_DMNconn <- readRDS(file = "../report/fig/figDMNconn.rds")
fig_DMNconn
```


## Exploratory Analyses: Effects of average and change BMI on FC
### Reward network
#### Model 2.1 and 2.2 (adjusting for mFD)
```{r "reward model, mean, change BMI", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]

R11 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
R12a = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
anvmeanBMI_adjmFD=anova(R11,R12a)
```

```{r reward model_BMIFD results, caption="Model 2a and 2b for aggregated reward network FC", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R1, R11),
  custom.coef.names = c("Intercept", "average BMI", "change BMI", "Age", "Sex", "logmFD"),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```
#### Model EA 2.3 (adjusting for average and change in mFD)
```{r "reward model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
anvmeanFD=anova(R1,R03)
anvwithinFD=anova(R1,R03)

p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
```

```{r reward_BMI_splitFD results, caption="Model EA 2.3 for aggregated reward network FC", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R1), custom.coef.names=c("Intercept", "average BMI", "change BMI", "average mFD", "change mFD", "Age", "Sex"), custom.model.names = "R1", float.pos = "h!",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = TRUE)
```

### DMN
#### Model 2.1 and 2.2 (adjusting for mFD)
```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]

R11 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
R12a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
anvmeanBMI_adjmFD=anova(R11,R12a)
```

```{r DMN_BMIFD results, caption="Model 2.1 and 2.2 for aggregated DMN FC", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(
  list(R1, R11),
  custom.coef.names = c("Intercept", "average BMI", "change BMI", "Age", "Sex", "logmFD"),
  custom.model.names = c("R0", "R1"),
  float.pos = "h!",
  center = TRUE,
  digits = 3,
  leading.zero = FALSE,
  single.row = TRUE,
  include.adjrs = FALSE,
  include.bic = FALSE,
  table = FALSE,
  use.packages = TRUE
)
```
#### Model 2.3 (adjusting for average and change in mFD)
```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
anvmeanFD=anova(R1,R03)
anvwithinFD=anova(R1,R03)

p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
```

```{r DMN_BMI_splitFD results, caption="Model 2.3 for aggregated DMN FC", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R1), custom.coef.names=c("Intercept", "average BMI", "change BMI", "average mFD", "change mFD", "Age", "Sex"), custom.model.names = "R1", float.pos = "h!",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = TRUE)
```

\newpage
### Anatomical Labelling for models

Find in the detailed Output of the SPM Anatomy toolbox, version 2.2c. For Clusters without any labels, the toolbox was unable to assign any grey matter area. Models evaluated were model 2a estimating the effects of average BMI and change in BMI, adjusting for age and sex (Table \@ref(tab:Tab2a)), then model 2b where we additionally adjusted for logmFD (Table \@ref(tab:Tab2b)), and model 2c estimating the effects of average BMI and BMI variability as well as average logmFD and change in logmFD, adjusting for age and sex (Table \@ref(tab:Tab2cfirst) to Table \@ref(tab:Tab2clast)). Last, the FD model estimating the unique effects of average logmFD and change in logmFD when adjusting for age and sex was evaluated (Table \@ref(tab:TabFDfirst) and Table \@ref(tab:TabFDlast)).


```{r, include = FALSE}
ToolboxOutputList <- readRDS(file = "../report/tab/DetailedFCTableList.rds")
TableList_colname <- kableExtra::linebreak(c("Cluster","Number of voxels\nin cluster","\\% of cluster\nvolume assigned", "Hemisphere","Area","\\% of area overlap\nwith cluster"))

TableList <- list()
for (i in 1:length(ToolboxOutputList$EntryList)){
  TableList[[i]] <-
    ToolboxOutputList$EntryList[[i]] %>%
    knitr::kable(
      escape = FALSE,     # use font spec of latex with kableExtra
      col.names = TableList_colname,
      row.names = FALSE,
      format = "latex",
      booktabs = T,
      linesep = "",      # disable "\\addlinespace" at every 5th line
      #label = name_model[i],
      caption = ToolboxOutputList$CaptionList[i],
    ) %>%
    kableExtra::kable_styling(latex_options = c('scale_down'))
}
```

```{r Tab2a, echo=FALSE}
i = 2
ToolboxOutputList$EntryList[[i]] %>%
  knitr::kable(
    escape = FALSE,
    col.names = TableList_colname,
    row.names = FALSE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = ToolboxOutputList$CaptionList[i],
  ) %>%
  kableExtra::kable_styling(latex_options = c('scale_down')) 
```

```{r "Tab2b", echo=FALSE}
i = 1
ToolboxOutputList$EntryList[[i]] %>%
  knitr::kable(
    escape = FALSE,
    col.names = TableList_colname,
    row.names = FALSE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = ToolboxOutputList$CaptionList[i],
  ) %>%
  kableExtra::kable_styling(latex_options = c('scale_down')) 
```

```{r "Tab2cfirst", echo=FALSE}
i = 3
ToolboxOutputList$EntryList[[i]] %>%
  knitr::kable(
    escape = FALSE,
    col.names = TableList_colname,
    row.names = FALSE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = ToolboxOutputList$CaptionList[i],
  ) %>%
  kableExtra::kable_styling(latex_options = c('scale_down')) 
```

```{r, echo=FALSE}
TableList[[4]]
TableList[[5]]
TableList[[6]]
```

```{r "Tab2clast", echo=FALSE}
i = 7
ToolboxOutputList$EntryList[[i]] %>%
  knitr::kable(
    escape = FALSE,
    col.names = TableList_colname,
    row.names = FALSE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = ToolboxOutputList$CaptionList[i],
  ) %>%
  kableExtra::kable_styling(latex_options = c('scale_down')) 
```

```{r "TabFDfirst", echo=FALSE}
i = 8
ToolboxOutputList$EntryList[[i]] %>%
  knitr::kable(
    escape = FALSE,
    col.names = TableList_colname,
    row.names = FALSE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = ToolboxOutputList$CaptionList[i],
  ) %>%
  kableExtra::kable_styling(latex_options = c('scale_down')) 
```

```{r "TabFDlast", echo=FALSE}
i = 8
ToolboxOutputList$EntryList[[i]] %>%
  knitr::kable(
    escape = FALSE,
    col.names = TableList_colname,
    row.names = FALSE,
    format = "latex",
    booktabs = T,
    linesep = "",
    caption = ToolboxOutputList$CaptionList[i],
  ) %>%
  kableExtra::kable_styling(latex_options = c('scale_down'))
    
```